---
title: "Final Project"
author: "Danielle Jones & Katie Perez"
format: html
editor: visual
  highlight-style: pygments
syntax: tango
execute: 
  echo: true
  message: false
  warning: false
fig-width: 9
fig-height: 7
bibliography: references.bib
link-citations: true
---

# Load libraries

```{r}
library(here)
library(rio)
library(tidyverse)
library(janitor)
library(ggthemes)
library(gghighlight)
library(tidyr)

```

# Load data

```{r}
here("data", "Olympic_Swimming_Results_1912to2020.csv")

olyswim <- read_csv(here("data", "Olympic_Swimming_Results_1912to2020.csv")) %>%
  characterize() %>%
  clean_names()
```

# Cleaning/organizing data

```{r}
olyswim <- olyswim %>%
  filter(year >= 1969, year <=2020,
         relay == 0, rank != 0, rank !=4, rank !=5) %>%
  select(-location, -relay) %>%
    group_by(team) %>%
  filter(team == "USA" | team == "CHN" | team == "FRA") %>%

  mutate(foo = str_replace_all(results, "\\.", ":")) %>%
  separate_wider_delim(foo, ":", names = c("h", "m", "s", "ms"), too_few = "align_end") %>%
  mutate(across(c(h:ms), as.numeric),
         across(c(h:ms), ~as.numeric(ifelse(is.na(.), "0", .))),
         results_sec = case_when(
           s == 0 ~ as.character(ms),
           s!= 0 ~ paste0((h * 360) + (m * 60) + (s), ".", ms)
         ),
         results_sec = as.numeric(results_sec))
  
```

# Creating gender and stroke dataframe

```{r}
olyswimgenderstroke <- olyswim %>%
  group_by(gender, stroke) %>%
  summarise(mean_time = mean(results_sec))

summarize(olyswimgenderstroke)
```

# Creating wide dataframe

```{r}
olyswim_genderwide <- olyswim %>%
  pivot_wider(
  names_from = gender, 
  values_from = athlete
  )
```





## Attempting to pivot_longer. I didn't see a way to pivot the data we are working with, so I chose to make a subset (small_team_data)and make it messy and then pivot_longer. I have achieved messy, but now I'm having trouble cleaning the mess. Insights and suggestings are welcome.


```{r}
small_team_data <- olyswim %>% 
  filter( distance_in_meters =="100m"|gender == "women") %>% 
  select(year, stroke, team, rank) 

wide_data <- pivot_wider(small_team_data, names_from = team, values_from = rank)

long_data <- pivot_longer(wide_data, cols == "USA"| cols == "ROC", names_to = "time", values_to = "value")
```




